State of the Practice for Medical Imaging Software

### 2022-07-11 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Fixed an instance of passive voice
Updates to include CarverEtAl2022 and SmithAndMichalski2022

\usepackage{hyperref}
%\usepackage{tikz}
%\usetikzlibrary{mindmap}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}

@@

To gain insights, we devised 10 research questions, which can be applied to MI,
as well as to other domains, of research software \citep{SmithEtAl2021}.  The
questions are designed to learn about the community's interest in, and
experience with, artifacts, tools, principles, processes, methodologies and
qualities.  When we mention artifacts we mean the documents, scripts and code
that constitutes a software development project. Example artifacts include
requirements, specifications, user manuals, unit tests, system tests, usability
tests, build scripts, API (Application Programming Interface) documentation,
READMEs, license documents, process documents, and code.  Once we have learned
what MI developers do, we then put this information in context by contrasting MI
software against the trends shown by developers in other research software
communities.  Our aim is to collect enough information to understand the current
pain points experienced by the MI software development community so that we can
make some preliminary recommendations for future improvements.

@@

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{Sec_AddressConcerns})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}).
	(Section~\ref{ch_recommendations})

@@

We designed a general methodology to assess the state of the practice for SC
software \citep{SmithEtAl2021}. Details can be found in
Section~\ref{ch_methods}.  Our methodology has been applied to MI software
\citep{Dong2021} and Lattice Boltzmann Solvers \citep{Michalski2021}.  This
methodology builds off prior work to assess the state of the practice for such
domains as Geographic Information Systems \citep{smith2018state}, Mesh
Generators \citep{smith2016state}, Seismology software
\citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. A person month is considered to be $20$ working
days ($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours
per day, or $20 \times 8 = 160$ person hours.

@@

We developed a methodology for evaluating the state of the practice of research
software \citep{SmithEtAl2021}.  The methodology can be instantiated for a
specific domain of scientific software, which in the current case is medical
imaging software for visualization.  Our methodology involves and engages a
domain expert partner throughout, as discussed in
Section~\ref{sec_vet_software_list}.  The four main steps of the methodology
are:

@@

As part of filling in the measurement template, we use freeware tools to collect
repository related data. \href{https://github.com/tomgi/git_stats}{GitStats}
\citep{Gieniusz2019} is used to measure the number of binary files as well as
the number of added and deleted lines in a repository. This tool is also used to
measure the number of commits over different intervals of time.
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}
\citep{Boyter2021} is used to measure the number of text based files as well as
the number of total, code, comment, and blank lines in a repository.

@@

Our interviews were guided by a list of 20 questions, which can be found in
\citet{SmithEtAl2021}. Some questions are about the background of the software,
the development teams, the interviewees, and how they organize the projects. We
also ask about the developer's understandings of the users. Some questions focus
on the current and past difficulties, and the solutions the team has found, or
will try. We also discuss the importance and current situations of
documentation. A few questions are about specific software qualities, such as
maintainability, understandability, usability, and reproducibility. The
interviews are semi-structured based on the question list; we ask follow-up
questions when necessary.  The interview process presented here was approved by
the McMaster University Research Ethics Board under the application number 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}.

@@

Software tools are used to support the development, verification, maintenance,
and evolution of software, software processes, and artifacts \citep[p.\
501]{GhezziEtAl2003}. MI software uses tools for CI/CD, user support, version
control, documentation and project management.  To answer
\rqref{RQ_CompareToolsProjMngmnt} we summarize the tool usage in these
categories, and compare this to the usage by the research software community.

As mentioned in Section~\ref{sec_result_correctness_verifiability}, we
identified five projects using CI/CD tools (about 17\% of the assessed
projects). CI/CD using projects were found by examining the documentation and
source code of all projects. The count of CI/CD usage may actually be higher,
since traces of CI/CD usage may not always appear in a repository.  This was the
case for a study of LBM software, where interviews with developers showed that
more projects used CI/CD than was evident from repository artifacts alone
\citep{Michalski2021}.  The 17\% utilization for MI software contrasts with the
high frequency with which continuous integration is recommended in research
software development guidelines \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}. Although there is currently little data available
on CI/CD utilization for research software, our impression is that CI/CD is not
yet common practice, despite its recommendation.  For LBM software at least the
situation is similar to MI software, with only 12.5\% of 24 LBM packages showing
evidence of CI/CD in their repositories \citep{Michalski2021}.  \wss{add
reference to CarverEtAl2022, once it is published.}

@@

The common use of version control for MI software illustrates considerable
improvement from the poor adoption of version control tools that Wilson lamented
in 2006 \citep{Wilson2006}.  The proliferation of version control tools for MI
matches the increase in the broader research software community.  A little over
10 years ago version control was estimated to be used in only 50\% of research
software projects \citep{Nguyen-HoanEtAl2010}, but even at that time
\citet{Nguyen-HoanEtAl2010} noted an increase from previous usage levels. A
survey in 2018 shows 81\% of developers use a version control system
\citep{AlNoamanyAndBorghi2018}. \citet{Smith2018} has similar results, showing
version control usage for alive projects in mesh generation, geographic
information systems and statistical software for psychiatry increasing from
75\%, 89\% and 17\% (respectively) to 100\%, 95\% and 100\% (respectively) over
a four year period ending in 2018. (For completeness the same study showed a
decrease in version control usage for seismology software over the same time
period, from 41\% down to 36\%).  Almost every software guide cited in
Section~\ref{Sec_CompareArtifacts} includes the advice to use version control.
The high usage of version control tools in MI software matches the trend in
research software in general.

@@

As Table~\ref{Tbl_Guidelines} suggests, an emphasis on documentation, especially
for new developers, is echoed in research software guidelines. Multiple
guidelines recommend a document explaining how to contribute to a project, often
named CONTRIBUTING. Tutorials, user guides and quick start examples are also
recommended. \citet{SmithAndRoscoe2018} suggests including instructions
specifically for on-boarding new developers. For open source software in general
(not just research software), \citet{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos and screenshots.

@@

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer two research questions: i) What
are the pain points for developers working on research software projects?
(\rqref{RQ_PainPoints}); and, ii) How do the pain points of developers from MI
compare to the pain points for research software in general?
(\rqref{RQ_ComparePainPoints}).  We go through each of the identified pain
points and include citations that contrast the MI experience with observations
from researchers in other domains.  Potential ways to address the pain points
are covered in Sections~\ref{Sec_AddressConcerns} and~\ref{ch_recommendations}.

Table \ref{tab_obstacles} shows the number of times the interviewees mentioned
the current and past obstacles in their projects.   We group these pain points
into three major categories of obstacles: \textit{resource}, \textit{balance},
and \textit{testing}. We put the less mentioned ones into the category
\textit{Others}. The interviewees may have provided multiple
answers to each question. Thus, when we summarize the frequency of different
responses, the total is sometimes larger than nine.

\begin{table}[!ht]
\centering
\begin{tabular}{llll}
\toprule
\multirow{2}{*}{Category} & \multirow{2}{*}{Obstacle} & \multicolumn{2}{c}{Num
 ans.} \\ \cline{3-4} &  & current & past \\ 
\midrule
\multirow{2}{*}{Resource} & Lack of fundings & 3 &  \\
 & Lack of time to devote to the project & 2 & 1 \\ 
\midrule
\multirow{3}{*}{Balance} & Hard to keep up with changes in OS and libraries & 1
&  \\
 & Hard to support multiple OS & 2 &  \\
 & Hard to support lower-end computers & 1 & 2 \\ 
\midrule
Testing & Lack of access to real-world datasets for testing & 3 & 2 \\ 
\midrule
\multirow{7}{*}{Others}
 & Hard to have a high level roadmap from the start & 1 &  \\
 & Not enough participants for usability tests & 1 &  \\
 & Only a few people fully understand the large codebase & 1 &  \\
 & Hard to transfer to new technologies & & 2 \\
 & Hard to understand users' needs & & 1 \\
 & Hard to maintain good documentations & & 1 \\ 
\bottomrule
\end{tabular}
\caption{\label{tab_obstacles}Current and past obstacles by the numbers of interviewees with the answers}
\end{table}

@@

	\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
	Development Time} A developer of pyLBM noted that their small development
	team has a lack of time to implement new features. Small development teams
	are common for LBM software packages (as shown in the measurement table
	excerpt in Figure~\ref{measurement_template_image}). Lack of time is also
	highlighted as a pain point by other research software developers
	\citet{PintoEtAl2018, PintoEtAl2016, WieseEtAl2019}.

Developers identified a resource related pain point caused by a lack of funding
and time.  Potential and proven solutions suggested by the interviewees include:

\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Licensing the software to commercial companies that integrate it into
their products;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for CI/CD to save time.
\end{itemize}

@@

With respect to balance, developers expressed difficulty balancing between four
factors: cross-platform compatibility, convenience to development \&
maintenance, performance, and security.  The potential and proven solutions are:

@@

For the pain point category of testing developers identified the the lack of
access to real-world datasets.  One pain point in the development process is the
lack of access to real-world datasets for testing. The developers' strategies to
address this are summarized in Section~\ref{Sec_AddressConcerns} \ad{to be
fixed}. One threat to correctness is: with huge datasets for testing, the tests
are expensive and time-consuming. Three interviewees endorsed self tests /
automated tests, which may save time for testing. Some proposed and proven
solutions are as follows:

@@

\subsection{Discussions on Software Qualities} \label{sec_interview_software_qualities}

Questions 15--18, and 20 cover the software qualities of correctness,
maintainability, understandability, usability, and reproducibility,
respectively. We discuss each quality below.

\begin{description}
\item[Q15.] Was it hard to ensure the correctness of the software? If there were
any obstacles, what methods have been considered or practiced to improve the
situation? If practiced, did it work?
\end{description}

@@

\begin{description}
\item[Q16.] When designing the software, did you consider the ease of future
changes? For example, will it be hard to change the system’s structure, modules,
or code blocks? What measures have been taken to ensure the ease of future
changes and maintains?
\end{description}

\begin{description}
\item[Q17.] Provide instances where users have misunderstood the software. What,
if any, actions were taken to address understandability issues?
\end{description}

@@

\begin{description}
\item[Q18.] What, if any, actions were taken to address usability issues?
\end{description}

@@

\begin{description}
\item[Q20.] Do you have any concern that your computational results won't be
reproducible in the future? Have you taken any steps to ensure reproducibility?
\end{description}

@@

\section{Lessons from MI Developers} \label{Sec_AddressConcerns}

The best practices from MI developers are taken to address the pain points
mentioned in Section~\ref{painpoints}.  We found the practices as part of the
qualitative data from developer interviews
(Section~\ref{sec_interview_methods}).  The practices we summarize can
potentially be emulated by the MI software packages that do not currently follow
them.  Moreover, these practices may also provide examples that can be followed
by other research software domains.

@@

@@

% Overview of scientific computing
We define Scientific Computing (SC) as ``the use of computer tools to analyze or
simulate mathematical models of real world systems of engineering or scientific
importance so that we can better understand and predict the system's behaviour''
\citep{Smith2006}. Many researchers consider SC as the third pillar of science
and engineering, along with theory and experiment \citep{Landau2005}. Almost all
areas in science and engineering use computers for modelling \citep{Golub2014},
and software plays an essential role in modern scientific research
\citep{Hannay2009, Wilson2014}.  Software development in SC depends on three
fields of knowledge: engineering or scientific domain knowledge, mathematical
algorithm knowledge, and computational algorithm knowledge \citep{Landau2005,
Mehta2015}. Thus, most SC software developers are scientists in SC domains
\citep{Wilson2014}. However, they do not always use modern software development
techniques, tools, and methods \citep{Wilson2014}. Therefore, we developed a
methodology for assessing the state of the practice for SC software
\citep{SmithEtAl2021}. We apply this process to Medical Imaging (MI) software
that belongs to a specific domain of SC.

@@

- add more on comparison to existing literature surveys
- ask Mike:
    - for journal version ask for more figures?
    - for journal double blind solution
    - can he submit
    - potential reviewers
    - feedback
- Radiology - should not exceed 6000 words, 50 refs, 8 figures and 4 tables

>>>>

\usepackage{hyperref}
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

%\usepackage{tikz}
%\usetikzlibrary{mindmap}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{arydshln}
\usepackage{paralist}
\usepackage{booktabs}

@@

To gain insights, we devised 10 research questions, which can be applied to MI,
as well as to other domains, of research software \citep{SmithEtAl2021,
SmithAndMichalski2022}.  The questions are designed to learn about the
community's interest in, and experience with, artifacts, tools, principles,
processes, methodologies and qualities.  When we mention artifacts we mean the
documents, scripts and code that constitutes a software development project.
Example artifacts include requirements, specifications, user manuals, unit
tests, system tests, usability tests, build scripts, API (Application
Programming Interface) documentation, READMEs, license documents, process
documents, and code.  Once we have learned what MI developers do, we then put
this information in context by contrasting MI software against the trends shown
by developers in other research software communities.  Our aim is to collect
enough information to understand the current pain points experienced by the MI
software development community so that we can make some preliminary
recommendations for future improvements.

@@

\begin{enumerate}
	\item[RQ\refstepcounter{rqnum}\therqnum \label{RQ_WhatProjects}:] What MI
	software projects exist, with the constraint that the source code must be
	available for all identified projects? (Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_HighestQuality}:] Which
	of the projects identified in \rqref{RQ_WhatProjects} follow current best
	practices, based on evidence found by experimenting with the software and
	searching the artifacts available in each project's repository?
	(Section~\ref{ch_results})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareHQ2Popular}:] How
	similar is the list of top projects identified in \rqref{RQ_HighestQuality}
	to the most popular projects, as viewed by the scientific community?
	(Section~\ref{Sec_VsCommunityRanking})
    \item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareArtifacts}:] How
	do MI projects compare to research software in general with respect to the
	artifacts present in their repositories?
	(Section~\ref{Sec_CompareArtifacts})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareToolsProjMngmnt}:]
	How do MI projects compare to research software in general with respect to
	the use of tools (Section~\ref{Sec_CompareTools}) for:
	\begin{enumerate} 
		\item [\rqref{RQ_CompareToolsProjMngmnt}.a] development; and,
		\item [\rqref{RQ_CompareToolsProjMngmnt}.b] project management?
	\end{enumerate}
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_CompareMethodologies}:]
	How do MI projects compare to research software in general with respect to
	principles, processes and methodologies used?
	(Section~\ref{Sec_CompareMethodologies})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_PainPoints}:] What are
	the pain points for developers working on MI software projects?
	(Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_ComparePainPoints}:] How
	do the pain points of developers from MI compare to the pain points
	for research software in general? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Concerns}:] For MI
	developers what specific best practices are taken to address the pain points
	and software quality concerns? (Section~\ref{painpoints})
	\item [RQ\refstepcounter{rqnum}\therqnum \label{RQ_Recommend}:]
	What research software development practice could potentially address the
	pain point concerns identified in \rqref{RQ_PainPoints}).
	(Section~\ref{ch_recommendations})

@@

We designed a general methodology to assess the state of the practice for SC
software \citep{SmithEtAl2021, SmithAndMichalski2022}. Details can be found in
Section~\ref{ch_methods}.  Our methodology has been applied to MI software
\citep{Dong2021} and Lattice Boltzmann Solvers \citep{Michalski2021}.  This
methodology builds off prior work to assess the state of the practice for such
domains as Geographic Information Systems \citep{smith2018state}, Mesh
Generators \citep{smith2016state}, Seismology software
\citep{Smith2018Seismology}, and Statistical software for psychology
\citep{smith2018statistical}.  In keeping with the previous methodology, we have
maintained the constraint that the work load for measuring a given domain should
be feasible for a team as small as one person, and for a short time, ideally
around a person month of effort. A person month is considered to be $20$ working
days ($4$ weeks in a month, with $5$ days of work per week) at $8$ person hours
per day, or $20 \times 8 = 160$ person hours.

@@

We developed a methodology for evaluating the state of the practice of research
software \citep{SmithEtAl2021, SmithAndMichalski2022}.  The methodology can be
instantiated for a specific domain of scientific software, which in the current
case is medical imaging software for visualization.  Our methodology involves
and engages a domain expert partner throughout, as discussed in
Section~\ref{sec_vet_software_list}.  The four main steps of the methodology
are:

@@

As part of filling in the measurement template, we use freeware tools to collect
repository related data. \href{https://github.com/tomgi/git_stats}{GitStats}
\citep{Gieniusz2019} is used to measure the number of binary files as well as
the number of added and deleted lines in a repository. We also use this tool to
measure the number of commits over different intervals of time.
\href{https://github.com/boyter/scc}{Sloc Cloc and Code (scc)}
\citep{Boyter2021} is used to measure the number of text based files as well as
the number of total, code, comment, and blank lines in a repository.

@@

We based uur interviews on a list of 20 questions, which can be found in
\citet{SmithEtAl2021}. Some questions are about the background of the software,
the development teams, the interviewees, and how they organize the projects. We
also ask about the developer's understandings of the users. Some questions focus
on the current and past difficulties, and the solutions the team has found, or
will try. We also discuss the importance and current situations of
documentation. A few questions are about specific software qualities, such as
maintainability, understandability, usability, and reproducibility. The
interviews are semi-structured based on the question list; we ask follow-up
questions when necessary.  The interview process presented here was approved by
the McMaster University Research Ethics Board under the application number 
\href{https://github.com/smiths/AIMSS/blob/master/StateOfPractice/MACREM/Application.pdf}
{MREB\#: 5219}.

@@

Software tools are used to support the development, verification, maintenance,
and evolution of software, software processes, and artifacts \citep[p.\
501]{GhezziEtAl2003}. MI software uses tools for CI/CD, user support, version
control, documentation, and project management.  To answer
\rqref{RQ_CompareToolsProjMngmnt} we summarize the tool usage in these
categories, and compare this to the usage by the research software community.

@@

The common use of version control for MI software illustrates considerable
improvement from the poor adoption of version control tools that Wilson lamented
in 2006 \citep{Wilson2006}.  The proliferation of version control tools for MI
matches the increase in the broader research software community.  A little over
10 years ago \citet{Nguyen-HoanEtAl2010} estimated that version control was used
in only 50\% of research software projects, but even at that time
\citet{Nguyen-HoanEtAl2010} noted an increase from previous usage levels. A
survey in 2018 shows 81\% of developers use a version control system
\citep{AlNoamanyAndBorghi2018}. \citet{Smith2018} has similar results, showing
version control usage for alive projects in mesh generation, geographic
information systems and statistical software for psychiatry increasing from
75\%, 89\% and 17\% (respectively) to 100\%, 95\% and 100\% (respectively) over
a four-year period ending in 2018. (For completeness the same study showed a
decrease in version control usage for seismology software over the same time
period, from 41\% down to 36\%).  A recent survey by \citet{CarverEtAl2022}
shows version control use among practitioners at over 95\%, with 83/87 survey
respondents indicating that they use it. All but one of the software guides
cited in Section~\ref{Sec_CompareArtifacts} includes the advice to use version
control. (The USGS guide \citep{USGS2019} was the only set of recommendations to
not mention version control.) The high usage of version control tools in MI
software matches the trend in research software in general.

As mentioned in Section~\ref{sec_result_correctness_verifiability}, we
identified five projects using CI/CD tools (about 17\% of the assessed
projects). We found CI/CD using projects by examining the documentation and
source code of all projects. The count of CI/CD usage may actually be higher,
since traces of CI/CD usage may not always appear in a repository.  This was the
case for a study of LBM software, where interviews with developers showed that
more projects used CI/CD than was evident from repository artifacts alone
\citep{Michalski2021}.  The 17\% utilization for MI software contrasts with the
high frequency with which research software development guidelines recommend
continuous integration \citep{BrettEtAl2021, Brown2015, ThielEtAl2020,
Zadka2018, vanGompelEtAl2016}. Although there is currently little data available
on CI/CD utilization for research software, our impression is that CI/CD is not
yet common practice, despite its recommendation.  For LBM software at least the
situation is similar to MI software, with only 12.5\% of 24 LBM packages showing
evidence of CI/CD in their repositories \citep{Michalski2021}.  The survey of
\citet{CarverEtAl2022} suggests higher use of CI/CD with 54\% (54/100)
respondents indicating that they use it.

@@

As Table~\ref{Tbl_Guidelines} suggests, an emphasis on documentation, especially
for new developers, is echoed in research software guidelines. Multiple
guidelines recommend a document explaining how to contribute to a project, often
named CONTRIBUTING. Guidelins also recommend tutorials, user guides and quick start examples. \citet{SmithAndRoscoe2018} suggests including instructions
specifically for on-boarding new developers. For open source software in general
(not just research software), \citet{Fogel2005} recommends providing tutorial
style examples, developer guidelines, demos, and screenshots.

@@

Based on interviews with nine developers (described in
Section~\ref{sec_interview_methods}), we answer three research questions:
\rqref{RQ_PainPoints}) What are the pain points for developers working on
research software projects?; \rqref{RQ_ComparePainPoints}) How do the pain
points of developers from MI compare to the pain points for research software in
general?; and \rqref{RQ_Concerns}) For MI developers what specific best practices
are taken to address the pain points and software quality concerns?  We go
through each of the identified pain points and include citations that contrast
the MI experience with observations from researchers in other domains.  We also
cover potential ways to address the pain points, as promoted by the community in
this section, and based on our experience in Section~\ref{ch_recommendations}.

Ways to address pain points found by community as part of the qualitative data
from developer interviews (Section~\ref{sec_interview_methods}).  The practices
we summarize can potentially be emulated by the MI software packages that do not
currently follow them.  Moreover, these practices may also provide examples that
can be followed by other research software domains.

\begin{itemize}
\item Hard to have a high level roadmap from the start
\item Not enough participants for usability tests
\item Only a few people fully understand the large codebase
\item Hard to transfer to new technologies
\item Hard to understand users' needs
\item Hard to maintain good documentations
\end{itemize}

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackDevTime}:] \textbf{Lack of
Development Time:} A developer of pyLBM noted that their small development team
has a lack of time to implement new features. Small development teams are common
for LBM software packages (as shown in the measurement table excerpt in
Figure~\ref{measurement_template_image}). Other domains of research software
also experience the lack of time pain point \citep{PintoEtAl2018, PintoEtAl2016,
WieseEtAl2019}. Developers identified a resource related pain point caused by a
lack of funding and time.  Potential and proven solutions suggested by the
interviewees include:

\begin{itemize}
\item Shifting from development to maintenance when the team does not have
enough developers for building new features and fixing bugs at the same time;
\item Improving documentation to save time answering users' and developers'
questions;
\item Supporting third-party plugins and extensions; and,
\item Using GitHub Actions for CI/CD to save time.
\end{itemize}

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_LackFunding}:] \textbf{Lack of
Incentive and Funding:} The TCLB developer noted a lack of incentives and
funding in academia for developing widely used research software. This problem
has also been noted by others \citep{gewaltig2012quality, Goble2014,
KaterbowAndFeulner2018}.  \citet{WieseEtAl2019} reported developer pains related
to publicity, since publishing norms have historically made it difficult to get
credit for creating software.  As studied by \citet{HowisonAndBullard2016},
research software (specifically biology software, but the trend likely applies
to other research software domains) is infrequently cited. \citet{PintoEtAl2018}
also mentions the lack of formal reward system for research software.

\begin{itemize}
\item Licensing the software to commercial companies that integrate it into
their products;
\end{itemize}

\item[P\refstepcounter{pnum}\thepnum \label{P_TechnologyHurdles}:]
\textbf{Technology Hurdles:} Technology pain points for LBM developers include
setting up parallelization and CI. The pain point survey of
\citet{WieseEtAl2019} also highlighted technical-related problems like
dependency management, cross-platform compatibility, CI, hardware issues and
operating system issues. Things mentioned by developers in interviews: Hard to
keep up with changes in OS and libraries, Hard to support multiple OS, and Hard
to support lower-end computers. With respect to balance, developers expressed difficulty balancing between four
factors: cross-platform compatibility, convenience to development \&
maintenance, performance, and security.  The potential and proven solutions are:

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_Correctness}:]
\textbf{Ensuring Correctness:} Several developers noted difficulties with
ensuring correctness. They alluded to difficulty with testing the correctness of
large numbers of features, and challenges with automated testing. The TCLB
developer commented that the amount of testing data needed was sometimes
problematic, since free testing services do not offer adequate facilities for
large amounts of data, requiring in-house testing solutions. Other research
software domains point to the following problems with testing: i)
\citet{PintoEtAl2018} mention the problem of insufficient testing; ii)
\citet{HannayEtAl2009} show that more developers think testing is important than
the number that believe they have a sufficient understanding of testing
concepts; and, iii) \citet{HannayEtAl2009, KanewalaAndBieman2013, KellyEtAl2011,
WieseEtAl2019} point to the oracle problem, which occurs in research software
when we do not have a means to judge the correctness of the calculated
solutions. Based on the interviews, the LBM experience seems to overlap with i
and ii; the developers did not allude to problem iii (the oracle problem).  As
discussed in Section~\ref{SecReproducibility}, the oracle problem likely did not
come up because the LBM developers have strategies for verifying their work.

For the pain point category of testing developers identified the lack of access
to real-world datasets.  One pain point in the development process is the lack
of access to real-world datasets for testing. The developers' strategies to
address this are summarized in this section. One threat to correctness is: with
huge datasets for testing, the tests are expensive and time-consuming. Three
interviewees endorsed self tests / automated tests, which may save time for
testing. Some proposed and proven solutions are as follows:

@@

@@

\item[P\refstepcounter{pnum}\thepnum \label{P_Usability}:]
\textbf{Usability:} Several developers noted that users sometimes try to use
incorrect LBM method combinations to solve their problems. Furthermore, some
users think that the packages will work out of the box to solve their cases,
while in reality CFD knowledge needs to be applied to correctly modify the
packages for the new endeavour.  Some respondents in the survey of
\citet{WieseEtAl2019} also mentioned that users do not always have the expertise
required to install or use the software.

@@

@@

blurb to explain the ones that are coming were not explicitly identified as pain
points.

\begin{enumerate}
\item \textbf{Maintainability:} The developer of ESPResSo said that they
originally wrote their source code with a specific application in mind, which
later caused too much coupling between components in the source code. This
resulted in technical debt \citep{KruchtenEtAl2012}, which has an impact on
future maintainability and reusability. Concern with technical debt is likely
why researchers in the survey of \citet{Nguyen-HoanEtAl2010} rate
maintainability as the third most important software quality. More recently the
push for sustainable software \citep{deSouzaEtAl2019} is motivated by the pain
that past developers have had with accumulating too much technical debt.

\item \textbf{Reproducibility:}  Interviewees stressed the importance of
documentation for both users and developers throughout the interviews. They
emphasize that a lack of time and funding (\ppref{P_LackDevTime}) has a negative
effect on the documentation. Most of the developers are scientific researchers
evaluated on the scientific papers that they produce. Writing and updating
documentation is something that is done in their free time, if that time arises.
Others also mention inadequate research software documentation
\citep{PintoEtAl2018, WieseEtAl2019}.  The problem also arises with non-research
software \citep{LethbridgeEtAl2003}. Recommendations on the development of
research software state that developers should critically evaluate their own
development processes in terms of quality assurance and comply with
international standards for software documentation
\citep{KaterbowAndFeulner2018}.

@@

@@

- add in ACs, borrow from previous paper.

@@

% Overview of scientific computing
We define Scientific Computing (SC) as ``the use of computer tools to analyze or
simulate mathematical models of real world systems of engineering or scientific
importance so that we can better understand and predict the system's behaviour''
\citep{Smith2006}. Many researchers consider SC as the third pillar of science
and engineering, along with theory and experiment \citep{Landau2005}. Almost all
areas in science and engineering use computers for modelling \citep{Golub2014},
and software plays an essential role in modern scientific research
\citep{Hannay2009, Wilson2014}.  Software development in SC depends on three
fields of knowledge: engineering or scientific domain knowledge, mathematical
algorithm knowledge, and computational algorithm knowledge \citep{Landau2005,
Mehta2015}. Thus, most SC software developers are scientists in SC domains
\citep{Wilson2014}. However, they do not always use modern software development
techniques, tools, and methods \citep{Wilson2014}. Therefore, we developed a
methodology for assessing the state of the practice for SC software
\citep{SmithEtAl2021, SmithAndMichalski2022}. We apply this process to Medical
Imaging (MI) software that belongs to a specific domain of SC.

@@

- CarverEtAl2022 has discussion of requirements etc. that we could also contrast (seems that people overestimate their use of requirements.)
- add more on comparison to existing literature surveys
- ask Mike:
    - for journal version ask for more figures?
    - for journal double blind solution
    - can he submit
    - potential reviewers
    - feedback
- Radiology - should not exceed 6000 words, 50 refs, 8 figures and 4 tables
